---
title: "Report: initial spatial-disease model"
author: "Mengru Yuan"
date: "August 20, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Long-term objective

Identify disease patterns, estimate the strength of the relationship between chronic diseases, and provide evidence for implementing potential public health interventions. 

## Objective for this report
Get to know the data by fitting the initial model using INLA package.   
The initial model is:  
$$
Y_{i,j} \sim Poisson (\lambda_{i,j})
$$
$$
log(\lambda_{i,j}) = log(n_{i,j}) + \beta_0 + \gamma_i + \delta_j
$$

$Y_{i,j}$: number of incidence cases in 2014 for chronic disease $j$ in region $i$  
$n_{i,j}$: person-year at risk in 2014 for chronic disease $j$ in region $i$  
$i$: $i_{th}$ CLSC region, from 1:57  
$j$: $j_{th}$ chronic disease, from 1:8  

$$
\gamma_i \sim Normal(0, \sigma^2)
$$
$$
\delta_j \sim Normal(0, \tau^2)
$$

## Data
The data include the incidence cases, person-year at risk for eight chronic diseases in 57 CLSC regions in Montreal CMA in 2014.   
The eight chronic diseases are ami(Acute Myocardial Infarction), asthma, chf(Congestive Heart Failure), copd (Chronic Obstructive Pulmonary Disease), diabetes, hypertension, ihd(Ischemic heart disease), stroke

```{r, echo=FALSE, include=FALSE}
library(INLA)
library(dplyr)
library(ggplot2)
```

### Dataset
```{r cars}
clsc.dis = read.table(file = "Data/disease_clsc_incidence_2014.csv", head=TRUE, sep = ',')
clsc.dis = clsc.dis[, -1]
clsc.dis = clsc.dis %>%
    group_by(disease, clsc) %>%
    summarise(cases = sum(num), pop = sum(denom))

str(clsc.dis)
clsc.dis$dis_fct = as.numeric(clsc.dis$disease)
clsc.dis$clsc_fct = as.numeric(as.factor(clsc.dis$clsc))
```

## Built the model with random effect on disease and clsc region
```{r}
formula.0 = cases ~ 1 +
    f(dis_fct, model="iid") +
    f(clsc_fct, model ="iid")
inla.model.0 = inla(formula.0, family="poisson", data=clsc.dis, offset=log(pop),
                    control.compute=list(cpo=TRUE,dic=TRUE),
                    control.predictor=list(compute=TRUE))


```

#### function for plot
```{r}
tbl_for_plot = function(inla_model, n, n_var){
    df = matrix(NA, 1000, n)
    for(i in 1:n){
        df [,i] <- inla.rmarginal(1000,marg = inla_model$marginals.random[[n_var]][[i]])}
    df.quartiles <- exp(t(apply(df, MARGIN=2,
                                function(x) quantile( x, probs= c(0.025,0.5,0.975)))))
    
    data.test <- data.frame(varb = c(1:n),
                            value = df.quartiles[,2],
                            ui=df.quartiles[,3],
                            li=df.quartiles[,1])
    return(data.test)
}
```


## Results:
### Plot the CI for posterior distribution of the random effect of each disease
```{r pressure, echo=FALSE}
data = tbl_for_plot(inla.model.0, 8, 1)
ggplot(data, aes(varb, value)) +
    geom_point(size = 3, shape = 15) +
    geom_abline(intercept = 1, slope = 0, linetype = "dotted", size = 1) +
    geom_errorbar(aes(ymin = li, ymax = ui), size = 1, width = 0.3) +
    labs(x = "disease", y = "Exp_disease_effect", title = "Disease effect") +
    scale_x_discrete(name = "Diseases", limits = c(1:8), labels = unique(clsc.dis$disease)) +
    theme(axis.text = element_text(size =10)) 
```

### plot the CI for posterior distribution of the random effect of each clsc
```{r}
data = tbl_for_plot(inla.model.0, 57, 2)
data$clsc = unique(clsc.dis$clsc)
data = data[order(data$value), ]
data$y = c(1:57)

ggplot(data, aes(y, value)) +
    geom_point(size = 3, shape = 15) +
    geom_abline(intercept = 1, slope = 0, linetype = "dotted", size = 1) +
    geom_errorbar(aes(ymin = li, ymax = ui), size = 1, width = 0.3) +
    labs(x = "CLSC", y = "Exp_CLSC_effect", title = "Spatial effect") +
    scale_x_discrete(name = "CLSC", limits = c(1:57), labels = as.character(data$clsc)) +
    theme(axis.text = element_text(size =5), 
          axis.text.x = element_text(angle = 45, hjust = 1)) 
```

